---
title: "Problem Set #1"
author: "Experiments and Causality"
date: \today
output: pdf_document
---
<!--
Some guidelines for submitting problem sets in this course:

- Please submit a PDF document rather than a Word document or a Google document.
- Please put your name at the top of your problem set.
- If you'll be using `R` or `Python` code to calculate your answers, please put the code and its output directly into your Problem Set PDF document.
- It is highly recommended, although not required, that you use the RMarkdown feature in RStudio to compose your problem set answers. RMarkdown allows you to easily intermingle analysis code and answers in one document. It is of a similar design as `jupyter` and an ipython notebook.
- You do not need to show work for trivial calculations, but showing work is always allowed.
- For answers that involve a narrative response, please feel free to describe the key concept directly and briefly, if you can do so, and do not feel pressure to go on at length.
- Please ask us questions about the problem set if you get stuck. **Don’t spend more than 20 minutes puzzling over what a problem means.** 
- Please ensure that someone (us!) can compile your solution set. The best way is to use the web-hosted links we've provided. 
- For text answers, please **embolden** your answers to the questions.
- For computing answers, please store them in the answer objects we have created. 
-->

# 1. Potential Outcomes Notation 

- Explain the notation $Y_{i}(1)$. **Ans:** This notation corresponds to the treatment outcome of the subject i or in other words it is the outcome if subject i is exposed to the treatment.  

- Explain the notation $E[Y_{i}(1)|d_{i}=0]$. **Ans:** It is the expectation of $Y_{i}(1)$ when one subject is selected at random from the subjects that were not treated. 

- Explain the difference between the notation $E[Y_{i}(1)]$ and the notation $E[Y_{i}(1)|d_{i}=1]$.
(Extra credit). **Ans:** $E[Y_{i}(1)]$ is the expectation of $Y_{i}(1)$ when one subject is selected at random from among all the subjects, while $E[Y_{i}(1)|d_{i}=1]$ is the expectation of $Y_{i}(1)$ when one subject is selected at random from the subjects that were treated.

- Explain the difference between the notation $E[Y_{i}(1)|d_{i}=1]$ and the notation 
$E[Y_{i}(1)|D_{i}=1]$.  Use exercise 2.7 from FE to give a concrete example of the difference.
**Ans:** $E[Y_{i}(1)|d_{i}=1]$ it is the expectation of $Y_{i}(1)$ when one subject is selected at random from the subjects that were actually treated, while $E[Y_{i}(1)|D_{i}=1]$ it is the expectation of $Y_{i}(1)$ when one subject is selected at random from the subjects that are supposed hypothetically treated. From excersise 2.7 (actual) $E[Y_{i}(1)|d_{i}=1] = 22.5$ while (supposed) $E[Y_{i}(1)|D_{i}=1]=30$. The difference in notation is that $Y_{i}(1)|d_{i}=1$ is the treated potential outcome that is actually treated and $Y_{i}(1)|D_{i}=1$ is the treated potential outcome for a subject that would be treated under some hypothetical allocation of treatments. In the excersise 2.7 this applies when the author says "suppose" that an experimenter randomly selects villages 3 and 7 and put them in the treatment group, we could write this as $Y_{3}(1)|D_{3}=1$ and $Y_{i}(7)|D_{7}=1$ but actually villages 1 and 7 were in actually allocated to villages 1 and 7, hence we could write $Y_{3}(1)|d_{3}=1$ and $Y_{i}(7)|d_{7}=1$. 

# 2. Potential Outcomes Practice
Use the values in the following table to illustrate that $E[Y_{i}(1)] - E[Y_{i}(0)] = E[Y_{i}(1) - Y_{i}(0)]$.

+---------------+------------+------------+----------+
|               | $Y_{i}(0)$ | $Y_{i}(1)$ | $\tau_i$ |
+===============+============+============+==========+
|Individual 1   |   5        |  6         |  1       |                          
|               |            |            |          |                        
+---------------+------------+------------+----------+
|Individual 2   |   3        |  8         |  5       |                          
|               |            |            |          |                          
+---------------+------------+------------+----------+
|Individual 3   |   10       |  12        |  2       |                          
|               |            |            |          |                          
+---------------+------------+------------+----------+
|Individual 4   |   5        |  5         |  0       |                          
|               |            |            |          |                      
+---------------+------------+------------+----------+
|Individual 5   |  10        |  8         |  -2      |                          
|               |            |            |          |                      
+---------------+------------+------------+----------+

```{r}
# Uploading the control and treatment groups outcome
po_control <- c(5,3,10,5,10)
po_treatment <- c(6,8,12,5,8)
tau <- po_control - po_treatment
```
$E[Y_{i}(1)] - E[Y_{i}(0)]$ equals to `r mean(po_control) - mean(po_treatment)` while $E[Y_{i}(1) - Y_{i}(0)]$ equals to `r mean(tau)` and they are the same.  

# 3. Conditional Expectations

Consider the following table:

+---------------+------------+------------+----------+
|               | $Y_{i}(0)$ | $Y_{i}(1)$ | $\tau_i$ |
+===============+============+============+==========+
|Individual 1   |   10       |  15        |  5       |                          
|               |            |            |          |                        
+---------------+------------+------------+----------+
|Individual 2   |   15       |  15        |  0       |                          
|               |            |            |          |                          
+---------------+------------+------------+----------+
|Individual 3   |   20       |  30        |  10      |                          
|               |            |            |          |                          
+---------------+------------+------------+----------+
|Individual 4   |   20       |  15        |  -5      |                          
|               |            |            |          |                      
+---------------+------------+------------+----------+
|Individual 5   |  10        |  20        |  10      |                          
|               |            |            |          |                      
+---------------+------------+------------+----------+
|Individual 6   |  15        |  15        |  0       |                           
|               |            |            |          |                      
+---------------+------------+------------+----------+
|Individual 7   |  15        |  30        |  15      |                          
|               |            |            |          |                      
+---------------+------------+------------+----------+
|Average        |  15        |  20        |  5       |                          
|               |            |            |          |                      
+---------------+------------+------------+----------+

Use the values depicted in the table above to complete the table below. 

```{r}
# Some calculations I will need for the next table.  
data <- data.frame(y0 = c(10,15,20,20,10,15,15), y1 = c(15,15,30,15,20,15,30))
```


$E[Y_{i}(0)|Y_{i}(1) > 15]$
```{r}
mean(data[data$y1>15,"y0"])
```
$E[Y_{i}(1)|Y_{i}(0) > 15]$
```{r}
mean(data[data$y0>15,"y1"])
```


+---------------+-------+-------+-------+--------------------------+
| $Y_{i}(0)$    | 15    | 20    | 30    | Marginal $Y_{i}(0)$      |
+===============+=======+=======+=======+==========================+
|10             | n:1   | n:1   |n:0    |                          |
|               | %:1/7 | %:1/7 | %:0   | 2/7                      |
+---------------+-------+-------+-------+--------------------------+
|15             | n:2   | n:0   | n:1   |                          |
|               | %:2/7 | %:0   | %:1/7 | 3/7                      |
+---------------+-------+-------+-------+--------------------------+
|20             | n:1   | n:0   | n:1   |                          |
|               | %:1/7 | %:0   | %:1/7 | 2/7                      |
+---------------+-------+-------+-------+--------------------------+
|Marginal       |       |       |       |                          |
| $Y_{i}(1)$    |  4/7  | 1/7   |  2/7  | 1.0                      |
+---------------+-------+-------+-------+--------------------------+

a. Fill in the number of observations in each of the nine cells; **Ans** See table above.  
b. Indicate the percentage of all subjects that fall into each of the nine cells. **Ans** See table above.  
c. At the bottom of the table, indicate the proportion of subjects falling into each category of $Y_{i}(1)$. **Ans** See table above.  
d. At the right of the table, indicate the proportion of subjects falling into each category of $Y_{i}(0)$. **Ans:** See table above.  
e. Use the table to calculate the conditional expectation that $E[Y_{i}(0)|Y_{i}(1) > 15]$. **Ans:** from the table 10/3 + 15/3 + 20/3 = 4 + 30/3 = 15 or  
$E[Y_{i}(0)|Y_{i}(1) > 15]$ = `r mean(data[data$y1>15,"y0"])`   
```{r}
mean(data[data$y1>15,"y0"])
```

f. Use the table to calculate the conditional expectation that $E[Y_{i}(1)|Y_{i}(0) > 15]$. **Ans:** From the table 15/2 + 30/2 = 22.5 or  
$E[Y_{i}(1)|Y_{i}(0) > 15]$ = `r mean(data[data$y0>15,"y1"])`  
```{r}
mean(data[data$y0>15,"y1"])
```



# 4. More Practice with Potential Outcomes
Suppose we are interested in the hypothesis that children playing outside leads them to have better eyesight.  

Consider the following population of ten representative children whose visual acuity we can measure.  (Visual acuity is the decimal version of the fraction given as output in standard eye exams.  Someone with 20/20 vision has acuity 1.0, while someone with 20/40 vision has acuity 0.5.  Numbers greater than 1.0 are possible for people with better than “normal” visual acuity.)

```{r, echo = F}
d <- data.frame(child = 1:10, 
                y0 = c(1.1, 0.1, 0.5, 0.9, 1.6, 2.0, 1.2, 0.7, 1.0, 1.1), 
                y1 = c(1.1, 0.6, 0.5, 0.9, 0.7, 2.0, 1.2, 0.7, 1.0, 1.1) )
knitr::kable(d)
```

In the table, state $Y_{i}(1)$ means “playing outside an average of at least 10 hours per week from age 3 to age 6,” and state $Y_{i}(0)$ means “playing outside an average of less than 10 hours per week from age 3 to age 6.”   $Y_{i}$ represents visual acuity measured at age 6.

a. Compute the individual treatment effect for each of the ten children.  Note that this is only possible because we are working with hypothetical potential outcomes; we could never have this much information with real-world data. (We encourage the use of computing tools on all problems, but please describe your work so that we can determine whether you are using the correct values.)

```{r, echo = F}
# The individual treatment effect is defined as the difference between two potential outcomes
answer.POa <- d$y1 - d$y0
d$tau <- answer.POa
d
```
a) **Ans:** See in the able above the tau showing the individual treatment effect.  

b. In a single paragraph, tell a story that could explain this distribution of treatment effects. **Ans:** There are no differences in visual acuity between children who play outside and those who don't.  

c. What might cause some children to have different treatment effects than others? **Ans:** This might be just a measurement error and we need to remember that this children are just one random sample of the children population.  

d. For this population, what is the true average treatment effect (ATE) of playing outside.

**Ans:**
```{r, echo = F}
answer.POd <- mean(answer.POa)
answer.POd
``` 

e. Suppose we are able to do an experiment in which we can control the amount of time that these children play outside for three years.  We happen to randomly assign the odd-numbered children to treatment and the even-numbered children to control.  What is the estimate of the ATE you would reach under this assignment? (Again, please describe your work.)

```{r, echo = F}
# Creating even and odd index
even_indexes<-seq(2,10,2)
odd_indexes<-seq(1,9,2)

# Simulating the assignment of odd-numbered children to treatment
po_treatment <- d[odd_indexes,]
# Simulating the assingment of even-numbered children to control
po_control <- d[even_indexes,]
# Calculating the estimate of the ATE
answer.POe <- mean(po_treatment$y1) - mean(po_control$y0)
# Here we are exposed to a selection bias because of the difference between the odd and even children.
answer.POe
``` 
** Ans:** Here the treatment effect is -0.06, slightly different from the true ATE.  


f. How different is the estimate from the truth?  Intuitively, why is there a difference?  

**Ans:** Here we might be exposed to a selection bias because if the researcher or any asssistant consiously decided to select the odd-numbered children to treatment group. If the researcher randomly assign the odd-numbered children to the treatment group assuming that he selected the odd-numbered children from all the possible alternatives randomly then this ATE of -0.06 would be fine otherwise not. If the researcher did not consider all possible options then we would be exposed to a selection bias. Hence, the difference with the true estimate `r -0.04 + 0.06` can be explained by the fact that we are no longer making a random selection on the treatment and control group so $E[Y_i(0)|D_i=1] - E[Y_i(0)|D_i=0]$ is no longer zero. This result can be explained by $E[Y_i(1)] - E[Y_i(0)] = E[Y_i(1) -Y_i(0)|D_i=1] + E[Y_i(0)|D_i=1] - E[Y_i(0)|D_i=0]$, where $E[Y_i(1) -Y_i(0)|D_i=1]=-0.18$ is the ATE among the treated and $E[Y_i(0)|D_i=1] - E[Y_i(0)|D_i=0]=0.06$ is the selection bias.  

See below calculation of  $E[Y_i(1) -Y_i(0)|D_i=1]=-0.18$ 

```{r}
# The ATE among the treated
mean(po_treatment$y0) - mean(po_control$y0)
```
To calculate $E[Y_i(0)|D_i=1] - E[Y_i(0)|D_i=0]=0.06$
```{r}
# The selection bias
mean(po_treatment$y1 - po_treatment$y0)
```

g. We just considered one way (odd-even) an experiment might split the children. How many different ways (every possible way) are there to split the children into a treatment versus a control group (assuming at least one person is always in the treatment group and at least one person is always in the control group)? **Ans:** There are $\sum_{k=1}^9 \binom{10}{k}$ 

```{r, echo = F}
s <- 0
for(i in 1:9) {
  # i-th element of `u1` squared into `i`-th position of `usq`
  s <- s + dim(combn(10,i))[2]
  paste(i,s)
}
paste("The number of possible ways to split is", s)

answer.POg <- s
```
**Ans:** There are 1022 posible ways to split the children in treatment and control group assuming that there would be at least one child per group.  

h. Suppose that we decide it is too hard to control the behavior of the children, so we do an observational study instead.  Children 1-5 choose to play an average of more than 10 hours per week from age 3 to age 6, while Children 6-10 play less than 10 hours per week.  Compute the difference in means from the resulting observational data.

```{r, echo = F}
# Creating 1-5 and 6-10 index
indexes1_5<-seq(1,5,1)
indexes6_10<-seq(6,10,1)

# Simulating the assignment of children 1-5 to treatment
po_treatment <- d[indexes1_5,]
# Simulating the assingment of children 6-10 to control
po_control <- d[indexes6_10,]
# Calculating the estimate of the ATE
answer.POh <- mean(po_treatment$y1) - mean(po_control$y0)
# Here we are exposed to a selection bias because because we are not manipulating "X". There might be a reson why 1-5 children, the ones who have the lowest vissual accuity choose to play more than 10 hours per week, maybe is just difficult for them to read and study so they prefere to play instead.
answer.POh
``` 
**Ans:** Since this is an observational study we are subject to the selection bias, which could explain this big difference between -0.44 and the true ATE = -0.04. 

i. Compare your answer in (h) to the true ATE.  Intuitively, what causes the difference?  

**Ans:** By just looking at the schedule of outcomes, I noticed that children 1-5 are the ones with poor visual acuity while the children from 6 to 10 are the ones with good visual acuity.  

<!--
# 5. FE, exercise 2.5

*Note that the book typically defines D to be 0 for control and 1 for treatment. However, it doesn't have to be 0/1. In particular, one can have more than two treatments, or a continuous treatment variable. Here, the authors want to define D to be the number of minutes the subject is asked to donate. (This is because “D” stands for “dosage.”)*
-->

# 5. Randomization and Experiments

Suppose that a reasearcher wants to investigate whether after-school math programs improve grades. The researcher randomly samples a group of students from an elementary school and then compare the grades between the group of students who are enrolled in an after-school math program to those who do not attend any such program. Is this an experiment or an observational study? Why? 

<!--
#7: Skip in 2017
# 7. FE, exercise 2.8
Peisakhin and Pinto report the results of an experiment in India designed to test the effectiveness of a policy called RTIA, which allows citizens to inquire about the status of a pending request from government officials. In their study, the researchers hired confederates, slum dwellers who sought to obtain ration cards (which permit the purchase of food at low cost). Applicants for such cards must fill out a form and have their residence and income verified. The table of results and conditions is below, and there is more in $FE$. 

+--------------------------+-------+------+-----+---------+
|                          | Bribe | RTIA | NGO | Control |
+==========================+=======+======+=====+=========+
| Number Confederates (C)  |    24 |   23 |  18 |      21 |
+--------------------------+-------+------+-----+---------+
| C w/ residence verif     |    24 |   23 |  18 |      20 |
+--------------------------+-------+------+-----+---------+
| M days to verif          |    17 |   37 |  37 |      37 |
+--------------------------+-------+------+-----+---------+
| C w/ ration card 365+    |    24 |   20 |   3 |       5 |
+--------------------------+-------+------+-----+---------+

a. Interpret the apparent effect of the treatments on the proportion of applicants who have their residence verified and the speed with which verification occurred. 
b. Interpret the apparent effect of the treatments on the proportion of applicants who actually received a ration card. 
c. What do these results seem to suggest about the effectiveness of the RTIA as a way of helping slum dwellers obtain ration cards? 
-->


# 6. Lotteries

A researcher wants to know how winning large sums of money in a national lottery affect people's views about the estate tax. The research interviews a random sample of adults and compares the attitudes of those who report winning more than $10,000 in the lottery to those who claim to have won little or nothing. The researcher reasons that the lottery choose winners at random, and therefore the amount that people report having won is random. 

a. Critically emvaluate this assumption. 
b. Suppose the researcher were to restrict the sample to people who had played the lottery at least once during the past year. Is it safe to assume that the potential outcomes of those who report winning more than $10,000 are identical, in expectation, to those who report winning little or nothing?  

a) **Ans:** I agree with the assumption that the researcher is selecting a random group of the population to ensure his research is representative of the entire adult population. Then his idea of creating the treatment group with people winning the lottery more than \$10,000 and the control group with people who won less than \$10,000 is correct because these two groups were created randomly and we can assume that there is no selection bias.

b) **Ans:** This second assumption will no longer create a true treatment effect applicable to the entired population. But since still there will be a random sampling of the people who won > \$10,000 (treatment group) and the people who played at least once during the past year and did not win more than \$10,000 (control group). Hence it will be applicable  $E[Y_{i}(0)|D=1] = E[Y_{i}(0)|D=0]$

*Clarifications* 

1. Please think of the outcome variable as an individual’s answer to the survey question “Are you in favor of raising the estate tax rate in the United States?”
2. The hint about potential outcomes could be rewritten as follows: Do you think those who won the lottery would have had the same views about the estate tax if they had actually not won it as those who actually did not win it? (That is, is $E[Y_{i}0|D=1] = E[Y_{i}0|D=0]$, comparing what would have happened to the actual winners, the $|D=1$ part, if they had not won, the $Y_{i}(0)$ part, and what actually happened to those who did not win, the $Y_{i}(0)|D=0$ part.) In general, it is just another way of asking, "are those who win the lottery and those who have not won the lottery comparable?"
3. Assume lottery winnings are always observed accurately and there are no concerns about under- or over-reporting.

# 7. Inmates and Reading
A researcher studying 1,000 prison inmates noticed that prisoners who spend at least 3 hours per day reading are less likely to have violent encounters with prison staff. The researcher recommends that all prisoners be required to spend at least three hours reading each day. Let $d_{i}$ be 0 when prisoners read less than three hours each day and 1 when they read more than three hours each day. Let $Y_{i}(0)$ be each prisoner's PO of violent encounters with prison staff when reading less than three hours per day, and let $Y_{i}(1)$ be their PO of violent encounters when reading more than three hours per day. 

In this study, nature has assigned a particular realization of $d_{i}$ to each subject. When assessing this study, why might one be hesitant to assume that ${E[Y_{i}(0)|D_{i}=0] = E[Y_{i}(0)|D_{i}=1]}$ and $E{[Y_{i}(1)|D_{i}=0] = E[Y_{i}(1)|D_{i}=1]}$? In your answer, give some intuitive explanation in English for what the mathematical expressions mean.

**Ans:** From the problem, maybe priosiners who are not violent are the ones who wants to read more than 3 hrs per day and maybe the violent prisioners are the one who doesn't want to read and read less than 3 hrs per day. Hence, the expectation of the potential outcome of a prisioner selected randomly within the treatment group will not be equal to the expectation of the potential outcome from a prisioner selected randomly from the whole group of prisioners. Therefore ${E[Y_i(1)|D_i=1] \neq E[Y_i(1)]}$. Since, we did not select randomly prisioners who actually read more than 3 hrs/day, we cannot expect that the expected potential outcome $Y_i(1)$ in the control group ${(D_i=0)}$ to be equal to the expected $Y_i(1)$ for the entire group of prisioners. Hence ${E[Y_i(1)|D_i=0] \neq E[Y_i(1)]}$. Thus, ${E[Y_i(1)|D_i=1] \neq E[Y_i(1)|D_i=0]}$. Applying similar analysis we can conclude that ${E[Y_i(0)|D_i=0] \neq E[Y_i(0)|D_i=1]}$ In summary those prisioners who read more than 3 hrs/day are not comparable to the prisioners who doesn't like to read.





